Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-03-25 doc/tfp-2011/main.tex
*** empty log message ***

\paragraph{Related Work.}
In order to compare and better distinguish
\of{Functory} from others work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item \textit{Distributed Functional Languages (DFLs)} --- functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
  Haskell, Erlang, etc.
\item \textit{Libraries for existing functional languages} --- that
  could be readily used in order to avoid implementing details like
  task distribution, fault-tolerance, socket programming, etc.
\end{enumerate}

@@

All the above mentioned languages could be used to implement our
library.  The main goal of \of{Functory}, though, is to provide the
users of an \emph{existing} functional programming language, high-level
user-friendly APIs that hide the messy details of task distribution
and fault-tolerance, enabling a ready-to-use library for efficient
utilization of computing resources.
We now turn to distributed computing libraries for general purpose
functional languages and weed out the distinguishing features of
\of{Functory} from other similar libraries.

@@

The main issue in any distributed computing environment is the ability
to handle faults, which is also a distinguishing feature of our
library.  
The fault tolerance mechanism of \functory\ is limited to workers;
handling master failures is the responsibility of the user, for
instance by periodically logging the master's state.
Worker faults are mainly of two kinds: either a worker is stopped,
and possibly later restarted; or a worker is temporarily or
permanently unreachable on the network. To provide fault tolerance,
our master implementation is keeping track of the status of each
worker.  This status is controlled by two timeout parameters $T_1$ and
$T_2$ and \of{Ping} and \of{Pong} messages sent by master and workers,
respectively. There are four possible statuses for a worker:
\begin{description}
\item[\of{not connected}:] there is no ongoing TCP connection between
  the master and the worker;
\item[\of{alive}:] the worker has sent some message
  within $T_1$ seconds;
\item[\of{pinged}:] the worker has not sent any message within $T_1$
  seconds and the master has sent the worker a
  \of{Ping} message within $T_2$ seconds;
\item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
  message (for more than $T_2$ seconds).
\end{description}
Whenever we receive a message from a worker, its status changes to
\of{alive} and its timeout value is reset.
\begin{center}
  \includegraphics{state.mps}
\end{center}

>>>>

\paragraph{Related Work.}
In order to compare and better distinguish
\of{Functory} from others work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item \textit{Distributed Functional Languages (DFLs)} --- functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
  Haskell, Erlang, etc.
\item \textit{Libraries for existing functional languages} --- that
  could be readily used in order to avoid implementing details like
  task distribution, fault-tolerance, socket programming, etc.
\end{enumerate}
\functory\ belongs to the second category.
For reasons of completeness, though, we first describe some existing DFLs
related to functional programming.

@@

Any DFL above could have been used to implement our library. 
Our motivation, though, was neither to implement our system using any
existing DFL nor to come up with a new DFL.
The goal of \of{Functory} is rather to provide the
users of an \emph{existing} general-purpose functional programming
language, namely \Ocaml, high-level 
user-friendly APIs that hide the messy details of task distribution
and fault-tolerance.
We now turn to distributed computing libraries for general purpose
functional languages and weed out the distinguishing features of
\of{Functory}.

@@

The main issue in any distributed computing environment is the ability
to handle faults, which is also a distinguishing feature of our
library.  
The fault tolerance mechanism of \functory\ is limited to workers;
handling master failures is the responsibility of the user, for
instance by periodically logging the master's state.
Worker faults are mainly of two kinds: either a worker is stopped,
and possibly later restarted; or a worker is temporarily or
permanently unreachable on the network. To provide fault tolerance,
our master implementation is keeping track of the status of each
worker.  This status is controlled by two timeout parameters $T_1$ and
$T_2$ and \of{Ping} and \of{Pong} messages sent by master and workers,
respectively. There are four possible statuses for a worker:
\begin{description}
\item[\of{not connected}:] there is no ongoing TCP connection between
  the master and the worker;
\item[\of{alive}:] the worker has sent some message
  within $T_1$ seconds;
\item[\of{pinged}:] the worker has not sent any message within $T_1$
  seconds and the master has sent the worker a
  \of{Ping} message within $T_2$ seconds;
\item[\of{unreachable}:] the worker has not yet responded to the \of{Ping}
  message (for more than $T_2$ seconds).
\end{description}
Whenever we receive a message from a worker, its status changes to
\of{alive} and its timeout value is reset.
\begin{center}
  \includegraphics[scale=0.7]{state.mps}
\end{center}
